The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-15 src/ms.tex
update learn section. Closes #42

\secauthor{Lía Corrales, David Shupe, Kelle Cruz + Learn team}

@@

\textit{Learn Astropy} is an umbrella term that acknowledges the broad
educational efforts made by the Astropy Project, which are led by the Learn
Team.
The efforts focus on developing online content and workshops covering
astronomy-specific coding tasks in Python.
As introduced in \citet{astropy:2018}, there are four different types of
Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook
lessons that are published in HTML format online; \textit{guides}, which are a
series of lessons providing a foundational resource for performing certain
type of astronomical analyses; \textit{examples}, which are snippets of code
that showcase a short task that can be performed with Astropy or an affiliated
package; and \textit{documentation}, which is contained within the code base.
This categorization drives content development, infrastructure choices, and the
appearance of the Learn Astropy website.
The Learn Team meets weekly to work on creating, expanding, improving these
educational resources.

\begin{description}

\item[Learn Astropy website] The Learn Team re-launched the main website and
search interface for Learn Astropy in 2021 with a new infrastructure platform,
built around full-text search and interactive filtering functionality, with the
goal of making content discoverable as the Learn Astropy content catalog
expands.
This work has been supported in part by a grant from the Dunlap Institute.
We have adopted Algolia, a search-as-a-service cloud platform, to store the
full-text and metadata records of Learn Astropy's content.
The new Learn Astropy website is a JavaScript (Gatsby/React) application that
uses the Algolia service to power its search and filtering user interface.
Our Python-based application, Learn Astropy Librarian, populates data into the
Algolia service. We tuned the Librarian around specific content formats (such as
Jupyter Notebook-based tutorial pages and Jupyter Book-based guides) to more
accurately index content and heuristically extract metadata.
A consequence of the new platform is that we now maintain and compile content
separately from the website application itself, enabling new content types.
Tutorials, which are written as Jupyter Notebooks, are now compiled into their
own Learn Astropy sub-site using
\href{https://github.com/astropy/nbcollection}{\texttt{nbcollection}}.
Guides, which utilize the \href{https://jupyterbook.org/}{Jupyter Book} build
infrastructure, are also deployed as separate websites using GitHub pages.
This architecture opens future possibilities of indexing third-party content,
hosted elsewhere, such as on institutional websites.

\item[Tutorials] Currently, we have \# of tutorials, spanning a wide range of astronomical topics and common tasks.

\item[Guides] {\it ccdproc} Jupyter book has been completed.

\item[Workshops] The Project has been conducting workshops at winter
meetings of the American Astronomical Society since AAS 225 in January 2015.
Up to the start of the coronavirus pandemic, these were full-day in-person
workshops with as many as 90 participants and a dozen facilitators from the
project.
During the pandemic, these workshops were moved to an online format and split
into basic and advanced sessions.
Additionally, beginning with the AAS 238 online meeting, the workshops have been
expanded to Summer AAS meetings.
The Learn team finds that the workshop audience is best found as AAS meetings as
opposed to more general Python meetings.

The Astropy Project provided another mode of community engagement at AAS
Meetings 235 and 237 by organizing a NumFOCUS Sponsored Projects booth in the
AAS Exhibit Hall.
Funding for the exhibit hall was provided alternately by NumFOCUS and later by
the Moore Foundation funding (\textbf{check this}).
The booth hosted a series of Q\&A special sessions during AAS 235 and webinars
during the virtual AAS 237 meetings, to provide the general astronomy community
information and access to experts on a variety of open source astronomical
tools.
\end{description}

@@

The Learn plan going forward is to continue to improve the tutorial website and
facilate new content;
to index third-party tutorials; and to look for supportable opportunities to
expand the reach of Astropy workshops beyond the American Astronomical Society
meetings.

>>>>

% \secauthor{Lía Corrales, David Shupe, Kelle Cruz + Learn team}

@@

\textit{Learn Astropy} is an umbrella term that acknowledges the broad
educational efforts made by the Astropy Project, which are led by the Learn
Astropy Team.
The efforts focus on developing online content and workshops covering
astronomy-specific coding tasks in Python.
As introduced in \citet{astropy:2018}, there are four different types of
Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook
lessons that are published in HTML format online; \textit{guides}, which are a
series of lessons providing a foundational resource for performing certain
type of astronomical analyses; \textit{examples}, which are snippets of code
that showcase a short task that can be performed with Astropy or an affiliated
package; and \textit{documentation}, which contains more detailed information
about the code base and user interface.
This categorization drives content development, infrastructure choices, and the
appearance of the \href{https://learn.astropy.org}{Learn Astropy website}.
The Learn Team meets weekly to work on creating, expanding, improving these
educational resources.

The Learn Team recently re-launched the main website and search interface for
Learn Astropy in 2021 with a new infrastructure platform, built around full-text
search and interactive filtering functionality, with the goal of making content
more easily searchable and discoverable as the Learn Astropy content catalog
expands.
This work has been supported in part by a grant from the Dunlap Institute.
We have adopted Algolia, a search-as-a-service cloud platform, to store the
full-text and metadata records of Learn Astropy's content.
The new Learn Astropy website is a JavaScript (Gatsby/React) application that
uses the Algolia service to power its search and filtering user interface.
Our \python-based application,
\href{https://github.com/astropy/learn-astropy-librarian/}{Learn Astropy
Librarian}, populates data into the Algolia service. We tuned the Librarian
around specific content formats (such as Jupyter Notebook-based tutorial pages
and Jupyter Book-based guides) to more accurately index content and
heuristically extract metadata.
A consequence of the new platform is that we now maintain and compile content
separately from the website application itself, enabling new content types.
Tutorials, which are written as Jupyter Notebooks, are now compiled into their
own Learn Astropy sub-site using
\href{https://github.com/astropy/nbcollection}{\texttt{nbcollection}}.
Guides, which utilize the \href{https://jupyterbook.org/}{Jupyter Book} build
infrastructure, are also deployed as separate websites using GitHub pages.
This architecture opens future possibilities of indexing third-party content,
hosted elsewhere, such as on institutional websites.

We currently host 19 tutorials written as interactive Jupyter Notebooks and
rendered into static HTML pages with the infrastructure described above.
The tutorials span a range of astronomical topics, from general tasks like
reading or creating FITS files with different content or working with
astronomical coordinate systems, to more specific exercises like analyzing
spectroscopic data from the UVES instrument.
Now that the backend infrastructure is stable, we are interested in collecting
new content to serve and share: If you have ideas for contributions, or have
tutorials hosted on third-party sites that you would like to make searchable
through the Learn Astropy interface, please reach out by creating a \github
issue in the \href{https://github.com/astropy/astropy-tutorials}{Astropy
Tutorials repository}.
We currently host one Guide --- a longer-form walkthrough of a more complex
concept or topic --- that is focused on
\href{http://www.astropy.org/ccd-reduction-and-photometry-guide}{CCD image
reduction and photometry}, but are also actively seeking new material that would
be suitable for new guides.

Beyond developing and serving educational content, the \astropy Project has been
conducting workshops at winter meetings of the American Astronomical Society
since AAS 225 in January 2015.
Up to the start of the coronavirus pandemic, these were full-day in-person
workshops with as many as 90 participants and a dozen facilitators from the
project.
During the pandemic, these workshops were moved to an online format and split
into basic and advanced sessions.
Additionally, beginning with the AAS 238 online meeting, the workshops have been
expanded to Summer AAS meetings.
The Learn team finds that the workshop audience is best found as AAS meetings as
opposed to more general \python meetings, as the content tends to be more
applicable for students and researchers in astronomy and astrophysics.

The Astropy Project recently provided another mode of community engagement at
AAS Meetings 235 and 237 by organizing a NumFOCUS Sponsored Projects booth in
the AAS Exhibit Hall.
Funding for the exhibit hall was provided alternately by NumFOCUS and later by
the Moore Foundation funding.
The booth hosted a series of Q\&A special sessions during AAS 235 and webinars
during the virtual AAS 237 meetings, to provide the general astronomy community
information and access to experts on a variety of open source astronomical
tools.

@@

The focus of Learn Astropy over the coming months and years is to further
improve the interface for sharing educational content on the main Learn Astropy
website, and to facilitate the development of new content that highlights
\astropy and \astropy-affiliated package functionality.
However, we simultaneously also plan to solicit indexing and ingestion of
third-party tutorial series, to provide a unified interface for identifying
astronomy-specific educational content that demonstrates the functionality
available in the vast ecosystem of open-source software packages.
We also plan to look for opportunities to expand the reach of \astropy workshops
beyond the AAS meetings.
